\LARGE \bf IMU-based pose-estimation for spherical robots with limited resources

### 2021-05-31 mfi21/ieeeconf/root.tex
small changes

To test the impact of the filter on 3D mapping, we used the prototype in Figure \ref{fig:prototype}.
After 2 s of initial standstill, the sphere rolled aprox. 4 m within 5 s in an indoor environment from one corner of the room to the opposing one.
We extracted the laserscan data, as well as the filtered pose using the presented algorithm.
Without further pre- or post-processing, the pose data is combined with the laserscan data to create a 3D map.
Figure \ref{fig:mapping} compares the result with another 3D map, which is generated in the same fashion.
However, a combination of the Madgwick filter for orientation and a double integration of the accelerometer mean for position, is used.
The presented filter clearly improves the resulting map compared to the Madgwick/integrator approach.
In particular, the walls and the ceiling are more recognizable, which is favourable for many SLAM algorithms.
Another improvement is that points were no longer sensed below the ground, i.e. the plane on which the sphere was moving on.
The more detailed part of the map, as seen in the bottom left corner of the above image of Figure \ref{fig:mapping}, corresponds to the initial standstill.
It is more detailed than the rest of the map because of the non-repeating flower-shaped scanning pattern, where point density increases with time.
Note that the same detailed part is also present in the image below.
However, it is out of view (at the bottom right corner) as the Madgwick filter misses the initial pose estimation by a lot.

>>>>

To test the impact of the filter on 3D mapping, we used the prototype in Figure \ref{fig:prototype}.
After 2 s of initial standstill, the sphere rolled aprox. 4 m within 5 s in an indoor environment from one corner of the room to the opposing one.
We extracted the laserscan data, as well as the filtered pose using the presented algorithm.
Without further pre- or post-processing, the pose data is combined with the laserscan data to create a 3D map.
Figure \ref{fig:mapping} compares the result with another 3D map, using the in the same run recorded pose-estimation by the Madgwick filter and simple rotation integration.
Therefore the same laserscan-data is mapped with different pose estimations.
The presented filter improves the resulting map compared to the Madgwick/integrator approach.
In particular, the walls and the ceiling are more recognizable, which is favourable for many SLAM algorithms.
Another improvement is that points were no longer sensed below the ground, i.e. the plane on which the sphere was moving on.
The more detailed part of the map, as seen in the bottom left corner of the above image of Figure \ref{fig:mapping}, corresponds to the initial standstill.
It is more detailed than the rest of the map because of the non-repeating flower-shaped scanning pattern, where point density increases with time.
Note that the same detailed part is also present in the image below.
However, it is out of view (at the bottom right corner) as the Madgwick filter misses the initial pose estimation by a lot.
